Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-01-27 docs/ACM/paper.tex
Definition of Windmill.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\columnwidth]{images/alg_main}
\caption{The main steps of our code generator.}
\label{fig:alg_main}
\end{center}
\end{figure}

@@

Definition~\ref{def:task} uses the concept of {\em memory region}.
Syntactically, memory regions are described by program variables and/or pointers
plus ranges of offsets that these pointers are allowed to dereference.
Given two tasks: $T_1 = (G_1, M_{i1}, M_{o1})$ and $T_2 = (G_2, M_{i2}, M_{o2})$,
if $M_{o1} \cap M_{i2} \neq \emptyset$, then we say that $T_2$ {\em depends} on
$T_1$.
If a program $P$ is partitioned into a set of $n$ tasks, then this partition is
said to be {\em correct} if it abides by Definition~\ref{def:correctness}.

\begin{definition}[Correctness]
\label{def:correctness}
A set $T$ of $n$ tasks is a correct parallelization of a program $P$ if:
\begin{compactitem}
\item $T$ does not contain cyclic dependence relations;
\item the execution of the tasks in $T$ in any ordering determined by dependence
relations, leads to the same results as the sequential execution of $P$.
\end{compactitem}
\end{definition}

\begin{example}[Memory Regions]
\label{ex:regions}
\fer{XX Add an example to describe the notions of regions and dependences.
The best is to modify one of the programs in Section 2, so that we can refer to
that program.
Pedro, can you think about it and propose something?}
\end{example}

@@

The {\Taskminer} algorithm is illustrated in FIgure \ref{fig:alg_main}. 
It uses information from both the \emph{Control Flow Graph} and
the \emph{Dependence Graph}. It relies on the concepts of
\emph{windmills} and \emph{helices} (Definitions \ref{def:helix} and \ref{def:windmill}).

\begin{definition}[Windmill]
\label{def:windmill}
Given a set $S$ of strongly connected nodes in the program \emph{Dependence Graph} $G$, 
a windmill $W$ is defined as:
$$W = \bigcup_i N_i[S]$$
\end{definition}

>>>>

\begin{figure}[b]
\begin{center}
\includegraphics[width=1\columnwidth]{images/alg_main}
\caption{The main steps of our code generator.}
\label{fig:alg_main}
\end{center}
\end{figure}

@@

Definition~\ref{def:task} uses the concept of {\em memory region}.
Syntactically, memory regions are described by program variables and/or pointers
plus ranges of dereferensable offsets.
Given two tasks: $T_1 = (G_1, M_{i1}, M_{o1})$ and $T_2 = (G_2, M_{i2}, M_{o2})$,
if $M_{o1} \cap M_{i2} \neq \emptyset$, then we say that $T_2$ {\em depends} on
$T_1$.
If a program $P$ is partitioned into a set of $n$ tasks, then this partition is
said to be {\em correct} if it abides by Definition~\ref{def:correctness}.

\begin{definition}[Correctness]
\label{def:correctness}
A set $T$ of $n$ tasks is a correct parallelization of a program $P$ if:
\begin{compactenum}
\item $T$ does not contain cyclic dependence relations;
\item the execution of the tasks in $T$ in any ordering determined by dependence
relations, leads to the same results as the sequential execution of $P$.
\end{compactenum}
\end{definition}

\begin{example}[Memory Regions]
\label{ex:regions}
Below we have two tasks, $T_{\mathit{foo}} = (\mathsf{foo}, \{\mathsf{v[i - 1]}\},
\{v[i]\})$ and $T_{\mathit{bar}} = (\mathsf{bar}, \{v[i]\}, \{\})$.
Dependencies are identified via the \textsf{depend} clause.
In this example, each SESE region is formed by one function call:
\end{example}

\includegraphics[width=1\columnwidth]{images/ex_depends}

@@

% Windmills and vanes
A {\em task candidate} is a program region that can run in parallel with the
rest of the program.
To identify task candidates, we rely on the structure of the program's
control flow graph.
To this end, we identify, subgraphs in the CFG called {\em windmills}, a term
coined by Rideau {\em et al.}~\cite{Rideau08} to describe structural relations
between register copies.
We adopt a slightly more general definition than Rideaus'; however, the
graphical shape of our graphs remain similar:

\begin{definition}[Windmill]
\label{def:windmill}
A windmill is a graph $G_w = G_c \cup G_{v1} \cup \ldots \cup G_{vn}$ formed by
a strongly connected component $G_c$, its center, plus $n$ components (not
necessarily strong) $G_{vi}$, the vanes, such that:
\begin{compactenum}
\item for any topological ordering of $G_w$, and nodes $n_c \in G_c$,
$n_v \in G_{vi}$, we have $n_c$ ahead of $n_v$;
\item for any $i$ and $j$, $1 \leq i < j \leq n$, $G_{vi}$ and $G_{vj}$ do not
share vertices.
\end{compactenum}
\end{definition}
